---
title: "Combining Data Frames Exercises"
format:
  revealjs: 
    chalkboard: true
    theme: serif
code-fold: true  
---

## Video [here](https://www.youtube.com/watch?v=st-MQY6pWeY)

```{r}
#| message: false
#| warning: false
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the dataset `happy_full`.

```{r}
#| echo: true
# install.packages("remotes")
# remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
# can also use devtools::install_github() since devtools is installed in
# system library
library(tidybiology)
data(happy)
data(happy_join_one)
data(happy_full)
```

## Problem One

* Take a look at the three data frames - `happy`, `happy_join_one`, and `happy_full`.  

```{r}
#| echo: fenced
#| label: "tbl-happy"
#| tbl-cap: "Happy Data"
happy |> 
  gt::gt()
```

##

See the results of `happy` in @tbl-happy, `happy_join_one` in @tbl-hjo, and the first six rows of `happy_full` in @tbl-hf.

##

```{r}
#| label: "tbl-hjo"
#| tbl-cap: "Happy Join One Data"
happy_join_one |> 
  gt::gt()
```

##

```{r}
#| label: "tbl-hf"
#| tbl-cap: "Happy Full Data"
happy_full |> 
  head() |> 
  gt::gt()
```

## Problem Two

* Combine `happy` and `happy_join_one` by row.

```{r}
#| echo: fenced
#| label: "tbl-br"
#| tbl-cap: "Result of using `bind_rows()` with `happy` and `happy_join_one`"
bind_rows(happy, happy_join_one) |> 
  gt::gt()
```
## Problem Three

* Combine `happy` and `happy_join_one` by column.  Note: Check to see if this is even possible.

```{r}
#| echo: fenced
#| label: "tbl-bc"
#| tbl-cap: "Result of using `bind_cols()` with `happy` and `happy_join_one`"
nrow(happy) == nrow(happy_join_one)
bind_cols(happy, happy_join_one) |> 
  gt::gt()
```
## Problem Four

* Which `join_*` function is a better alternative to the previous solution?

```{r}
#| echo: true
happy |> 
  inner_join(happy_join_one) |> 
  gt::gt()
```

## Problem Five

* Filter out all countries from `happy_full` that are also in `happy`.

```{r}
#| echo: TRUE
happy_full |> 
  filter(!country_name %in% happy$country_name) |> 
  DT::datatable()
```
## 

```{r}
#| echo: true
happy_full |> 
  anti_join(happy) |> 
  DT::datatable()
```

